---
title: "SQL"
output: html_document
---
## Filtering Columns and Rows

with SQL, you can filter columns with SELECT and rows with WHERE.  Lets try an expample with the Lahman baseball database. First, we need to load the appropriate packages. 
```{r warning=FALSE,message=FALSE}
library(Lahman)
library(sqldf)
```
Now, suppose we would like to see the homerun totals for the 1927 Yankees. We could write the following

```{r}
query<-"Select playerID,yearID,teamID,HR FROM Batting 
WHERE teamID='NYA' and yearID=1927"
sqldf(query)




```






```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting 
WHERE teamID= 'NYA' and HR>=40"

sqldf(query)
```

```{r}
query<-"SELECT playerID,teamID,HR,SO FROM Batting
WHERE HR>40 AND SO<60"

sqldf(query)
```

```{r}
query<- "SELECT playerID,teamID,yearID,HR FROM Batting
WHERE teamID= 'PHI' and HR>30 and yearID>=1970 and yearID<=1979 and HR>30"

sqldf(query)
```

## ORDER BY

With SQL you an order the rows of your output. FOr instance, suppose you want to see every instance of a player hitting more than 50 homeruns, but you would like the players with the most homeuns at the top. We could do the following:

```{r}
query<- "SELECT playerID,teamID,yearID,HR FROM Batting
WHERE HR>50
ORDER BY HR DESC"

sqldf(query)
```

This is my 7 pm class. I am still practicing. Let's make a query:

```{r}

query<-"SELECT playerID,HR FROM
Batting
WHERE playerID='ruthba01'"

sqldf(query)


```

